---
title: 'Copepods true colors: astaxanthin pigmentation as an indicator of individual fitness'
author: "Laure Vilgrain - March 2021"
output:
  html_document:
    code_folding: hide
    highlight: pygments
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float: yes
  df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, cache=FALSE, message = FALSE, warning = FALSE, comment="")
```

Line from R studio 

```{r include=FALSE}
library("tidyverse")
library("DT")

library("scales")
library("cowplot")
library("patchwork")
library("viridis")

library("rnaturalearth")
library("rnaturalearthdata")
library("sf")
library("sp")
library("rgeos")
library("s2")

library("car")

```

# 0. Read data 

**Suppl. Data SI. ** List of all the publications that discussed and/or quantified carotenoid pigmentation in copepods
```{r}
SI <- read_delim("SI_general_info.csv", ";", escape_double = FALSE, trim_ws = TRUE)
datatable(select(SI,-comments), rownames = FALSE, filter="top", options = list(pageLength = 5, scrollX=T))
```
<br/>
**Suppl. Data SII.** List of the detailed quantifications of carotenoids forms in copepods found in 38 articles from the literature
```{r}
SII <- read_delim("SII_asta_form_detailed.csv", ";", escape_double = FALSE, trim_ws = TRUE)
datatable(SII, rownames = FALSE, filter="top", options = list(pageLength = 5, scrollX=T))
```

<br/>
**Suppl. Data SIII.** Abiotic and biotic variables effects on copepod astaxanthin concentrations from the literature
```{r}
SIII <- read_delim("SIII_forcing_variables.csv", ";", escape_double = FALSE, trim_ws = TRUE)
datatable(SIII, rownames = FALSE, filter="top", options = list(pageLength = 5, scrollX=T))
```

<br/>
**Suppl. Data SIV.** List of the effects of carotenoid pigmentation on fitness-related variables in copepods
```{r}
SIV <- read_delim("SIV_fitness_effects.csv", ";", escape_double = FALSE, trim_ws = TRUE)
datatable(SIV, rownames = FALSE, filter="top", options = list(pageLength = 5, scrollX=T))
```

<br/>
<br/>

# I. Red pigmentation in copepods 

## A. Redness: a shared and conserved functional trait in copepods 

### *Studies*
```{r echo=T, results='hide'}
# Counts of all studies 
(SI %>% distinct(study, year, no) %>% nrow())
#-> 82

# Counts of studies with quantification of carotenoids
SI %>% filter(study_type !="no_quantification") %>% distinct(study, year, no)  %>% nrow()
#-> 75

# Ratio between both
(SI %>% filter(study_type !="no_quantification") %>% distinct(study, year, no)  %>% nrow()) / (SI %>% distinct(study, year, no) %>% nrow())
# -> 0.9146341
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

**82 articles** contain the words 'copepods” and “carotenoid”, “astaxanthin” or “redness” and among them, **75 (91%) quantified carotenoids** (in various units and methods). 

</div>



### *Genus and species*

```{r, echo=T, results='hide'}
# Number of species
SI %>% group_by(species) %>% summarise() %>% nrow()
# -> 157

# Number of genus
SI %>% filter(genus != "mixed") %>% group_by(genus) %>% summarise() %>% nrow()
# -> 62
```

```{r, fig.height=6, fig.width=5, fig.align="center", fig.cap=c("Figure S1. Counts of studies that observed carotenoid pigmentation in different copepod genus.")}
# Most studied genus 
SI %>% filter(genus != "mixed") %>% 
  ggplot() + 
  geom_bar(aes(y=fct_rev(fct_infreq(genus))))+
  labs(y="Genus", x= "Number of studies (all units)")+theme_linedraw(base_size = 10)
```


<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
Carotenoids were observed in **157 species among 61 genus** (we removed case where genus are mixed together, noted as "mixed"). We can see that *Diaptomus* and *Calanus* species are the most studied. 
</div>


<br/>
In some articles, carotenoid pigmentation was just discussed in the text of the article but not quantified, so we classify the study type as "no_quantification" in the "study_type" column. In other cases, pigmentation was assessed visually for some tests but no numerical quantifications were performed. In those cases we put NAs in the "concentration" column. **For all next analyses, only studies that quantified carotenoids will be used:**

```{r echo=T, results='hide'}
SI <- SI %>% filter(study_type !="no_quantification")
SI  %>% distinct(study, year, no, latitude, longitude) %>% nrow()
# -> 133
SI %>% filter(is.na(concentration) == FALSE) %>% distinct(study, year, no, latitude, longitude, altitude, species, stage) %>% nrow()
# -> 262
```


<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
Carotenoids quantifications were reported in **133 locations globally**, which correspond to **262 measurements of pigments**  (combinations study/latitude/longitude/altitude/species/stages) <br/>
</div>

<br/>

### *Units*

```{r}
SI  %>% filter(is.na(concentration) == FALSE) %>% group_by(unit) %>% summarise(nb_of_studies_using_this_unit = n()) %>% arrange(-nb_of_studies_using_this_unit) %>% ungroup()
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
Among the **262 quantifications**, the most common unit was **µg/mgDW**: 202 measurements, so only values in µg/mgDW will be kept for quantitative comparisons.
</div>
<br/>

### *Freshwater vs marine ecosystems*

```{r}
# Filter measurments in µg/mgDW
ug <- SI %>% filter (unit == "ug/mgDW")

# Transform to have numeric values 
ug$concentration <- as.numeric(ug$concentration)

# Count number of studies by ecosystems 
ug %>% group_by(ecosystem) %>% summarise(counts = n(), mean_concentration = mean(concentration), sd = sd(concentration)) %>% ungroup()
```


```{r echo=FALSE, fig.height=10, fig.width=16, fig.cap=c("Figure 1. Carotenoid concentrations in freshwater and marine copepods at the individual level across the world.")}

# Density distributions of pigment concentrations in freshwater and marine ecosystems: 
pigment_dd <- ug %>% 
  ggplot +
  #geom_histogram(aes(x = mean, fill = ecosystem, color = ecosystem), alpha =0.85)+
  geom_density(aes(x = concentration, fill = ecosystem, color = ecosystem), alpha =0.85)+
  #scale_x_continuous(trans='log10')+
  scale_x_log10(breaks = trans_breaks("log10",  function(x) 10^x),
             labels = trans_format("log10", math_format(10^.x))) +
  annotation_logticks(sides="b")+
  labs(x="Carotenoids concentration (μg/mgDW)",
       y="Density",
       fill= "Ecosystem")+
  theme_bw(base_size=22)+
  ylim(0,1.5)+
  theme(legend.position = "none", panel.grid.minor = element_blank())+
  scale_fill_manual(values=c("#7fc97f","#0868ac"))+
  scale_color_manual(values=c("#7fc97f","#0868ac"))+
  geom_vline(xintercept = mean(filter(ug, ecosystem == "marine")$concentration), linetype="dotted", 
                color = "#0868ac", size=1)+
  geom_vline(xintercept = mean(filter(ug, ecosystem == "freshwater")$concentration), linetype="dotted", 
                color = "#7fc97f", size=1)


# World map with locations of pigmented copepod observations: 

# Create a dataframe with locations of all studies (not only quantifications because observation is enough)
locations <- SI %>% distinct(study, year, no, latitude, longitude, ecosystem) %>% ungroup()

# Project the studies locations 
loc_proj <- st_as_sf(filter(locations, is.na(latitude) == FALSE), coords = c("longitude", "latitude"), 
    crs = 4032, agr = "constant")

# Find coastlines and borders
world <- ne_countries(scale = "medium", returnclass = "sf")

# Plot the map 
world_map <- ggplot(data = world) +
  geom_sf(size=0.05)+ #plot the world map 
  geom_sf(data = loc_proj, size = 3, alpha = 0.85, aes(color = ecosystem, shape = ecosystem))+ #project data points
  coord_sf(crs = "+proj=laea +lat_0=42 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs ")+ #spherical projection
  labs(color="Ecosystem", shape = "Ecosystem")+
  scale_color_manual(values=c( "#7fc97f", "#0868ac"))+
  theme_minimal(base_size=20)

ggdraw(pigment_dd) +  draw_plot(world_map, width =0.75, height =0.75, 
        x = -0.018, y = 0.25)



```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

Freshwater species have an average carotenoid concentration of **2.77 ± 2.71 µg.mgDW-1 (n=62)** and marine species an average carotenoid concentration of **0.44 ± 1.06 µg.mgDW-1 (n=140)**. 
</div>

<br/>
<br/>

We performed statisical tests to compare carotenoid mean concentrations between freshwater and marine ecosystems

```{r echo=T, results='hide'}
ug$ecosystem = factor(ug$ecosystem, levels = c("freshwater", "marine"))

# First, test of normal distribution of data 
shapiro.test(filter(ug, ecosystem =="freshwater")$concentration)
# -> W = 0.73354, p-value = 2.782e-09
shapiro.test(filter(ug, ecosystem =="marine")$concentration)
# -> W = 0.39528, p-value < 2.2e-16

```
-> concentrations are not normally distributed

```{r echo=T, results='hide'}
ug$concentration_log10 <- log10(ug$concentration)
# -> log transformation

shapiro.test(filter(ug, ecosystem =="freshwater")$concentration_log10)
# -> W = 0.98474, p-value = 0.6355 : normal distribution for freshwater data
shapiro.test(filter(ug, ecosystem =="marine")$concentration_log10)
# -> W = 0.9587, p-value = 0.0003215
# -> log-transformed marine concentrations are still not normal after transformation

leveneTest( log10(concentration) ~ ecosystem, data = ug, alternative = "two.sided" )
# -> and variances are not equal :
#        Df F value    Pr(>F)    
# group   1  15.819 9.739e-05 ***

```
-> log-transformed marine concentrations were **still not normal, and variances are not equal** 


```{r echo=T, results='hide'}
# Try to remove data from Fisher, suspected to be outliers ---
noFisher <- ug %>% filter(study != "Fisher et al.")
noFisher$concentration_log10 <- log10(noFisher$concentration)
shapiro.test(filter(noFisher, ecosystem =="freshwater")$concentration_log10)
# -> W = 0.98474, p-value = 0.6355 : normal distribution
shapiro.test(filter(noFisher, ecosystem =="marine")$concentration_log10)
# W = 0.98338, p-value = 0.6206 : normal distribution
leveneTest( log10(concentration) ~ ecosystem, data = noFisher, alternative = "two.sided" )
# -> but variances still not equal, Welch transformation needed
t.test(log10(concentration) ~ ecosystem, data = noFisher, alternative = "two.sided")
# -> t.test : freshwater and marine concentration are significantly different
```

-> we tried to **remove data from Fisher et al.**, suspected to be outliers. Distributions were then normal and freshwater and marine carotenoid concentrations **significantly different**. However, **we decided to keep all the complete dataset** with data from Fisher et al. (83 points more, many different species) and **use a non-parametric test:**

```{r}
# So we can use a non-parametric test: 
wilcox.test(filter(ug, ecosystem =="marine")$concentration_log10, filter(ug, ecosystem =="freshwater")$concentration_log10, alternative = "two.sided")
# -> significantly different 

```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
Freshwater and marine carotenoid concentrations are **significantly different** using a Wilcoxon rank sum test. We can then observe mean concentration differences between copepod genus:  
</div>


```{r echo=FALSE, fig.align="center", fig.cap=c("Figure S2. Carotenoid concentration (µg/mgDW) in different freshwater and marine copepod genus. Color represent the number of studies that were used to compute the mean carotenoid content."), fig.height=11, fig.width=14}
ug %>% filter(genus != "mixed") %>% 
  group_by(genus, ecosystem) %>% 
  summarise(nb_studies = n(), pigment_content = mean(concentration), sd = sd(concentration)) %>% 
  ungroup() %>% 
  ggplot() + 
  geom_bar(aes(pigment_content, genus, fill = nb_studies), stat ="identity")+
  facet_wrap(~ecosystem)+
  labs(x="Carotenoid content (µg/mgDW)", fill = "Number of studies (units in µg/mgDW only)")+
  theme_minimal(base_size = 16)
```


<br/>

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
Plenty of copepod genus have the availability to synthesize astaxanthin but their propensity to do so seems mediated by the environment where they live. Note that then marine species accumulating the more pigments are *Tigriopus sp.* living in shallow temperate waters (ex: tidal pools of rocky shores) and *Calanus spp.* living in arctic and subarctic waters, submitted to strong seasonal climate variations.  
</div>

<br/>

## B. Localisation, molecular forms and metabolism of carotenoids pigments

### *Copepods' carotenoids types (astaxanthin, canthaxanthin, etc)*

```{r results='hide'}
# Number of studies that quantified the different types of carotenoids in copepods
detailed_carotenoids <- SI %>% filter(!is.na(astaxanthin_on_total_car)) 
detailed_carotenoids %>% group_by(study, year, no) %>% summarise(obs = n()) %>% nrow()
# -> 30 studies 

asta_car <- SI %>% group_by(astaxanthin_on_total_car) %>% summarise(counts = n())
sum(asta_car$counts) # on a total of 291 observations
asta_car <- asta_car %>% filter(!is.na(astaxanthin_on_total_car)) # remove NA
sum(asta_car$counts) 
# -> 166 observations with detailed information about carotenoids and 125 observations without detailed information about carotenoids

# For these 30 studies and over 166 quantification, distinguishing  observations of astaxanthin in minority (<50%) or majority(>50%)
  # Characters strings : 
min1 <- filter(asta_car, astaxanthin_on_total_car == "minority")
dom1 <- filter(asta_car, astaxanthin_on_total_car == "dominance")

  # Numerical values (remove characters)
asta_car <- asta_car %>% filter(astaxanthin_on_total_car!= "minority" & astaxanthin_on_total_car != "dominance")
asta_car$astaxanthin_on_total_car <- as.numeric(asta_car$astaxanthin_on_total_car)
dom2<- asta_car %>% filter(astaxanthin_on_total_car >= 50)
min2 <- asta_car %>% filter(astaxanthin_on_total_car < 50)

  # Bind into dataframes in order to count
dom <- rbind(dom1, dom2)
min <- rbind(min1, min2)
sum(dom$counts) / 166 *100
sum(min$counts) / 166 *100
# -> in 96%, astaxanthin was dominant (and really often largly dominant : >90% of copepod carotenoids) - 160 quantification 
# -> in ~4%, astaxanthin was in a minority or <60% of carotenoids - 6 quantifications 

# Canthaxanthin : 
asta_cantha <- SI %>% filter(str_detect(comments, "canthaxanthin")) %>% select(study, year, no, comments) # if authors mentioned canthaxanthin in one observation, we always reported it in the column "comments" of the dataset
setdiff(asta_cantha$study, detailed_carotenoids$study) # coherently, all studies that mentioned canthaxantin (9) are studies that have detailed all the carotenoids of copepods (30)
nrow(asta_cantha %>% group_by(study, year, no) %>% summarise(n()))
nrow(asta_cantha)
nrow(asta_cantha) / 166 *100
# -> canthaxanthin was found in only 9 studies (21 observations, 13% of cases), and almost everytime in "negligible amounts, traces or <5%"
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
**Thirty studies** provide a detailed description of carotenoid pigments molecular forms for various species and developmental stages (166 measurements). In **96% (160 quantifications), astaxanthin was dominant** (or >50%), in ~5% (6 quantifications) of the cases, astaxanthin was in a minority or <50% of carotenoids. **Canthaxanthin was found in only 9 studies** (21 observations, 13% of cases), and almost every time in "negligible amounts, traces or <5%". 
</div>
<br/>

### *Astaxanthin forms*

```{r echo=T, results='hide'}
# Number of studies that detailed astaxanthin forms (free, mono and di-esters)
nrow(SII) # 38 detailed observations of astaxanthin forms
distinct(SII, study, year) # 18 studies

# Character strings : 
free_dom_est_min1 <- SII %>% filter(free == "dominant" & all_esters == "minority")
nrow(free_dom_est_min1) # Free form dominant and esters in minority: 7 obs
free_min_est_dom1 <- SII %>% filter(free == "minority" & all_esters == "dominant")
nrow(free_min_est_dom1) # Ester form dominant and free form in minority: 3 obs

# Remove charcter strings to have only numerical values
SII <- SII %>% filter(free != "dominant" & free != "minority")
SII$free <- as.numeric(SII$free)
SII$all_esters <- as.numeric(SII$all_esters)
free_dom_est_min2 <- SII %>% filter(free >= 50)
nrow(free_dom_est_min2) # Free form >= 50%: 16obs
free_min_est_dom2 <- SII %>% filter(free < 50)
nrow(free_min_est_dom2) # Free form < 50%: 12obs

free_dom_est_min <- rbind(free_dom_est_min1, free_dom_est_min2)
free_min_est_dom <- rbind(free_min_est_dom1, free_min_est_dom2 )

nrow(free_dom_est_min) / 38 * 100  
# -> in 61% of cases, free astaxanthin is dominant
nrow(free_min_est_dom)  / 38 * 100
# -> in 39% of cases astaxanthin esters are dominant

# Look at the proportions of mono- and di-esters
rbind(free_dom_est_min2, free_min_est_dom2) %>% 
  filter(!is.na(`mono-esters`)) %>% 
  select(-asta_form, - all_esters) %>% 
  gather(key = "form", value = "percent" , "free", `mono-esters`,`di-esters`) %>% 
  group_by(form) %>% 
  summarise(mean_percent = mean(percent))
# -> very detailed analyses (8 quantifications from 7 studies) show that on average, 53% of astaxanthin is in the free form, 27% of astaxanthin is in mono-esters and 20% is di-ester. 

```


<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
**18 studies** provided detailed forms of carotenoids (38 observations). In **61%** of cases, **free astaxanthin is dominant**, and in **39%** of cases **esters are dominant**. Very detailed analyses (8 quantifications from 7 studies) show that on average, 53% of astaxanthin is in the free form, 27% of astaxanthin is in mono-esters and 20% is di-ester.
</div>
<br/>

### *Ontogenic development*

This graphic illustrates that we don't have enough study that quantified carotenoid concentration along ontogenic development. We can not conclude about carotenoid quantities and forms according to copepod stages. <br/>

```{r echo=FALSE, fig.height=3, fig.width= 5, fig.align="center", fig.cap=c("Figure S3. Carotenoid concentrations in copepod stages from three different studies. Points are colored according to the unit used.")}
SI %>% filter((study == "Hairston" & no == "b")| study == "Holeton et al." | study == "Lotocka et al.") %>% mutate(Stage = factor(stage,levels = c("eggs", "nauplii", "copepodits", "adults", "adults_females"))) %>% ggplot() + geom_point(aes(Stage, concentration, color = unit, shape = study), size=3)+ labs(y="Carotenoid concentrations (units in color)", x="Development stage")+ theme_bw(base_size = 10)
```


# II. Effects of abiotic and biotic forcings on astaxanthin concentrations in copepods

## A. Environmental controls on pigmentation 

### *Latitude*

```{r fig.align="center", fig.cap=c("Figure S4. Carotenoid concentration (µg/mgDW, log-transformed) according to latitude (° in absolute value) in freshwater and marine ecosystems. Linear models are fitted but are not significant."), fig.height=3, fig.width=5}
ug  %>% 
  ggplot() + 
  geom_point(aes(abs(latitude), log10(concentration), color = ecosystem))+ 
  geom_smooth(aes(abs(latitude), log10(concentration), group = ecosystem, color = ecosystem), method = lm)+
  scale_color_manual(values=c("#7fc97f","#0868ac"))+
  theme_bw(base_size = 10)+
  labs(y= "Carotenoid content (µg/mgDW) - log ", x="Latitude (°absolute value)", color = "Ecosystem")

```

```{r}
test_lat_marine <- ug %>% filter(ecosystem == "marine") %>% mutate(lat_abs= abs(latitude))
summary(lm(log10(test_lat_marine$concentration)~test_lat_marine$lat_abs))

test_lat_freshwater <- ug %>% filter(ecosystem == "freshwater") %>% mutate(lat_abs= abs(latitude))
summary(lm(log10(test_lat_freshwater$concentration)~test_lat_freshwater$lat_abs))
```


<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
**Carotenoids content shows no relationship with latitude, neither in freshwater nor in marine environments.** In fact, concentration variations within a study (or a species) are often more important than variations according to latitude. </br>
</div>
<br/>


### *Altitude*

```{r fig.align="center", fig.cap=c("Figure S5. Carotenoid concentration (µg/mgDW, log-transformed) according to lake altitude (m) with a linear model fitted (significant, p-value < 0.001)."), fig.height=3, fig.width=5}

ug  %>% filter(ecosystem=="freshwater") %>% ggplot() + 
  geom_point(aes(altitude, log10(concentration)), color="grey40") + 
  geom_smooth(aes(altitude, log10(concentration)), method = "lm", color="black")+
  labs(x="Altitude (m)", y=" Carotenoid content (µg/mgDW) - log")+
  theme_bw(base_size = 10)

```
```{r}
summary(lm(log10(ug$concentration)~ug$altitude))
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
There is a **significant relationship between carotenoids content and altitude of lakes, p-value < 0.001**. However, we see a lot of noise (R square = 0.16) that could be explain by variations of other abiotic variables (temperature, lake depth, water turbidity) and biotic variables (phytoplanktonic community, predators, disponibility of other UV protectants).
</div>
<br/>


## B. Biological and metabolic drivers of pigmentation

### *Seasons*

```{r}
SI %>% filter(is.na(seasonal_pic_season) == FALSE) %>% select(study, year, no, seasonal_pic_season) %>% group_by(seasonal_pic_season) %>% summarise(count = n()) %>% mutate(tot = sum(count), percent = count/tot*100) %>% select(-tot)
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
On **37 time series**, astaxanthin content showed seasonal patterns with **maxima in winter (45%) or in spring (40%)** or in automn (3%). In 10% of the cases, no seasonal maxima were observed. 
</div>

<br/>


## C. Impacts of intra- and inter-species interactions on pigmentation

### *Predation in lakes*
```{r fig.align="center", fig.cap=c("Figure S6. Distribution of carotenoid concentrations (µg/mgDW, log-transformed) from lakes with and without predator presence. Carotenoid concentration difference according to predator presence is alomost significant (Wilcoxon test, p-value = 0.05014)"), fig.height=3.5, fig.width=5.5}
ug$predation = ifelse(ug$predator_abundance=="null", "no", "yes")

ug  %>% filter(ecosystem=="freshwater") %>%  group_by(predation) %>% summarise(n())

ug  %>% filter(ecosystem=="freshwater" & is.na(predation)== FALSE) %>% 
  ggplot() + 
  geom_boxplot(aes(predation, concentration)) + 
  geom_point(aes(predation, concentration, color=altitude)) +
  scale_color_viridis_c()+
  scale_y_continuous(trans='log10')+
  labs(y="Carotenoids content (ug/mgDW) 
       - log", color = "Altitude (m. asl)", title ="Predator presence: 33 yes, 18 no")+
  theme_bw(base_size = 10)


```


```{r echo=T, results='hide'}
#Test comparison predation/no predation (<30)
df_fish <- ug %>% filter(predation == "yes" | predation == "no") %>% mutate(predation = as.factor(predation))

#Normality ? 
shapiro.test(log10(filter(df_fish, predation =="yes")$concentration))
# W = 0.90082, p-value = 0.005596 -> normal distribution
shapiro.test(log10(filter(df_fish, predation =="no")$concentration))
# W = 0.92556, p-value = 0.09914 -> not normal distribution

```

Because n<30 and data not normaly distributed, we use a **non-parametric test** to compare carotenoid concentrations between lakes **with or without predators**:

```{r}
wilcox.test(concentration ~ predation, data = df_fish)
df_fish %>% group_by(predation) %>% summarise(mean(concentration))
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
We are **really close to the significance threshold (p-value = 0.05014)** for the Wilcocxon rank sum test to compare pigment content between lakes **with** (mean carotenoid concentration = **2.12 µg/mgDW, n=33**) and **without fish** (mean carotenoid concentration = **3.35 µg/mgDW, n = 22**). Altitude could play a role inside each category : high altitude lakes or ponds without fish can also be shallow and exposed to strong light radiations.
</div>

<br/>

### *Summary of all abiotic and biotic drivers*

```{r, fig.width=6, fig.height=8, fig.align='center', fig.cap=c("Figure 4. Counts of correlations between copepod carotenoid pigmentation and main environmental, biological or ecological forcings tested in the litterature.")}
# Change astaxanthin evolution column into factor
SIII$asta_correlation <- SIII$asta_correlation %>% str_replace_all(c("increase" = "positive", "no change" = "none", "decrease" = "negative"))
SIII$asta_correlation <- factor(SIII$asta_correlation, levels =c("positive","none", "negative"), ordered = TRUE)

# Replace names to have a more lisible figure 
SIII$forcing_variable <- SIII$forcing_variable %>% 
  str_replace_all(c("diet_composition" = "A. Diet quality", "feeding_activity" = "B. Feeding", "diet_quantity" = "C. Diet Quality", "fish_presence" = "D. Predator", "diurnal_variations" = "E. Daily cycle", "UV_radiations" = "F. UV", "natural_light" = "G. Natural light", "UV_fish" = "H. UV + Predator", "latitude_discrete" = "I. Latitude", "individual_depth" = "J. In situ depth", "ice_cover" = "K. Ice cover", "temperature" = "L. Temperature", "lake_elevation" = "M. Lake altitude", "dissolved_organic_carbon" = "N. DOC", "copper"= "O. Copper", "lake_depth" = "P. Lake depth", "egg_ratio"= "Q. Egg ratio", "body_length" = "R. Body length", "lipid_content" = "S. Lipid content", "MAAs" = "T. MAAs", "removal of UV radiations" = "U. UV removal"))
  
  
# Count the number of times a forcing variables was statistically tested according to carotenoid concentrations through a positive, negative or absent correlation:
SIII_grouped <- SIII %>%
  group_by(forcing_variable) %>% 
  mutate(n=n()) %>% 
  filter(n>=2) %>% 
  ungroup() %>% 
  group_by(forcing_variable, asta_correlation) %>% 
  summarise(counts=n()) %>% 
  ungroup()
  
SIII_grouped %>% 
  filter(forcing_variable != "U. UV removal" & forcing_variable != "UV_temp") %>% #because these 2 variables are to difficul to interpret without context 
  ggplot() + 
  facet_wrap(~forcing_variable, ncol = 4)+
  geom_bar(aes(x=asta_correlation, y=counts, fill=asta_correlation), stat="identity")+
  labs(x = "Types of correlation between various forcing 
    variables and individual carotenoid content:", y= "Counts of tested correlations in the litterature", fill = "Correlation")+
  scale_fill_manual(values=c("#238b45", "#91bfdb", "#990000"))+
  theme_bw(base_size = 12)+
  theme(strip.background = element_rect(colour="black", fill="white", linetype="solid"),
        axis.text.x = element_blank(), legend.position="bottom", legend.box = "horizontal",
         legend.title = element_blank(), legend.text = element_text(size = 12))

```


# III. How does redness impact copepod fitness (reproduction, growth, survival) ?

```{r}
# If we keep the whole dataset : 
SIV %>% 
  group_by(fitness_effect) %>%
  summarise(n=n(), percent = n()/(33+13+9))
# 9 negative (16%), 13 neutral (24%) and 33 positive effects (60%). 

```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
If we keep the whole dataset (SIV), we can find 9 negative (16%), 13 neutral (24%) and 33 positive effects (60%) of carotenoid pigmentation on copepod fitness. However, we decided to **remove data tagged with "low confidence"** (weak statistical tests or poor data quality discussed in the text of the article, see column "comments" of SIV for more details). 
</div>


<br/>

### *Counts of redness effects on various fitness variables*


```{r, fig.width=18, fig.height=10, fig.align='center', fig.cap=c("Figure 5. Carotenoid pigmentation effects on reproduction, growth and survival of copepods.")}

# Counts by fitness effect, arrange by fitness component (growth, survival or reproduction)
SIV_grouped <- SIV %>%
  filter(response_variable != "NA")%>%
  filter(confidence!= "low") %>% 
  group_by(fitness_effect, response_variable, graph_order,fitness_component) %>%
  summarise(n=n()) %>% 
  ungroup()%>%  
  spread(key=fitness_effect, value= n) %>% 
  arrange(-graph_order)

# Create a factor to have variables in a good order on the graphic (according to fitness component and discussion order in the body of the article)
SIV_grouped <- SIV_grouped %>%  mutate(index= as.factor(seq(1:nrow(SIV_grouped))))

# Positive ---
pos <- SIV_grouped %>% select(response_variable, positive, index, fitness_component) #keep only postive to have a color scale by effect type
pos$positive<- as.factor(pos$positive) # transform as factor to have the best color scale
green <- c("#41ae76", "#238b45", "#006d2c","#00441b") #because 4 levels

ppos <- ggplot(pos, aes(y = index, x= "", fill = positive))+
  geom_tile()+
  scale_fill_manual(values=green, na.value = "white", na.translate = F)+
  scale_y_discrete(labels=pos$response_variable)+
  labs(y="Fitness response variables", title = "Favorable - 31 cases (62%)")+
  theme_minimal(base_size = 15)+
  theme(legend.position="bottom",   axis.title.x = element_blank(), legend.title = element_blank(), legend.text = element_text(size = 15))

# Neutral ---
neu<- SIV_grouped %>% select(response_variable, neutral, index) #keep only postive to have a color scale by effect type
neu$neutral<- as.factor(neu$neutral) # transform as factor to have the best color scale
blue <- c("#a6bddb", "#74a9cf", "#3690c0") #because 3 levels

pneu <- ggplot(neu, aes(y = index, x = "", fill = neutral))+
  geom_tile()+
  scale_fill_manual(values=blue, na.value = "white", na.translate = F)+
  theme_minimal(base_size = 15)+
  labs(x= "Redness effects on copepod fitness - counts of tested observations:", title = "Neutral - 11 cases (22%)")+
  scale_y_discrete(labels=element_blank())+
  theme(legend.position="bottom", legend.title = element_blank(),  axis.title.y = element_blank(), legend.text = element_text(size = 15))

# Negative ---
neg<- SIV_grouped %>% select(response_variable, negative, index) #keep only postive to have a color scale by effect type
neg$negative <- as.factor(neg$negative) # transform as factor to have the best color scale
red <- c("#ef3b2c", "#cb181d","#a50f15") #because 3 levels

pneg <- ggplot(neg, aes(y = index, x = "", fill = negative))+
  geom_tile()+
  scale_fill_manual(values=red, na.value = "white", na.translate = F)+
  theme_minimal(base_size = 16)+
  scale_y_discrete(labels=element_blank())+
  labs(fill = "", title = "Unfavorable - 8 negative cases (16%)") +
  theme(legend.position="bottom", axis.title.y = element_blank(), axis.title.x = element_blank(), legend.title = element_blank(), legend.text = element_text(size = 18))


ppos | pneu | pneg
#ggsave("figures/fitness_redness.pdf", width=10, height =6)
```

```{r}
# However, we can remove points with "low" confidence, (bad statistics or data quality discussed in the text)
SIV %>% filter(confidence != "low") %>% 
  group_by(fitness_effect) %>%
  summarise(n=n(), percent = n()/(31+11+8))
# 8 negative effects (16%), 11 neutral (22%), 31 positive effects (62%)

# If we look in particular to survival and mortality curves : 
SIV %>% filter(str_detect(response_variable, "Survival") | str_detect(response_variable, "Mortality")) %>% 
  group_by(fitness_effect) %>%
  summarise(n=n(), percent = n()/21)
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
Out of **50 statistically tested** effects on one fitness variable, red pigmentation was **unfavorable for 8 (16%), neutral for 11 (22%), and favorable in 31 cases (62%)**. If we look in particular to survival and mortality curves, when copepods are in general submitted to an external stressor, we can see that carotenoid pigmentation is favorable in **71% of the cases**.
</div>






