---
title: "Copepods' true colors: astaxanthin pigmentation as an indicator of fitness"
author: "Laure Vilgrain - March 2021"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, cache=FALSE, message = FALSE, warning = FALSE, comment="")
```


```{r include=FALSE}
library("tidyverse")
library("DT")
library("writexl")

library("scales")
library("cowplot")
library("patchwork")
library("viridis")

library("rnaturalearth")
library("rnaturalearthdata")
library("sf")
library("sp")
library("rgeos")
library("s2")

library("car")

```

# 0. Read data 

**Suppl. Data SI.** List of all the publications that discussed and/or quantified carotenoid pigmentation in copepods
```{r}
SI <- read_delim("DataS1_general_info.csv", ";", escape_double = FALSE, trim_ws = TRUE)
SI
```
<br/>
**Suppl. Data SII.** List of the detailed quantifications of carotenoids forms in copepods found in 38 articles from the literature
```{r}
SII <- read_delim("DataS2_asta_form_detailed.csv", ";", escape_double = FALSE, trim_ws = TRUE)
SII
```

<br/>
**Suppl. Data SIII.** Abiotic and biotic variables effects on copepod astaxanthin concentrations from the literature
```{r}
SIII <- read_delim("DataS3_forcing_variables.csv", ";", escape_double = FALSE, trim_ws = TRUE)
SIII
```

<br/>
**Suppl. Data SIV.** List of the effects of carotenoid pigmentation on fitness-related variables in copepods
```{r}
SIV <- read_delim("DataS4_fitness_effects.csv", ";", escape_double = FALSE, trim_ws = TRUE)
SIV
```

<br/>
<br/>

# I. Carotenoid pigmentation in copepods 

## A. A shared and conserved functional trait in copepods 

### *Studies*
```{r results='hide'}
# Counts of all studies 
(SI %>% distinct(study, year, no) %>% nrow())
#-> 95

# Counts of studies with quantification of carotenoids
SI %>% filter(str_detect(study_type, "no_quantification", negate = T)) %>% distinct(study, year, no)  %>% nrow()
#-> 70

# Ratio between both
(SI %>% filter(str_detect(study_type, "no_quantification", negate = T)) %>% distinct(study, year, no)  %>% nrow()) / (SI %>% distinct(study, year, no) %>% nrow())
# -> 0.74
```



<div style="background-color: #bbe1e0">
**95 articles** contain the words 'copepods” and “carotenoid”, “astaxanthin” or “redness”, and have observed blue or red copepods in marine or freshwater ecosystems. Among them, **70 studies (74%) quantified carotenoids** (in various units and methods). 
</div>


### *Genus and species*

```{r, results='hide'}
# Number of species
SI %>% group_by(species) %>% summarise() %>% nrow()
# -> 171

# Number of genus
SI %>% filter(genus != "mixed") %>% group_by(genus) %>% summarise() %>% nrow()
# -> 65
```

```{r, fig.height=7.5, fig.width=5, fig.align="center", fig.cap=c("Counts of studies that observed carotenoid pigmentation in different copepod genus.")}
# Most studied genus 
SI %>% filter(genus != "mixed") %>% 
  ggplot() + 
  geom_bar(aes(y=fct_rev(fct_infreq(genus))))+
  labs(y="Genus", x= "Number of reported observations")+theme_linedraw(base_size = 10)
```


<div style="background-color: #bbe1e0">
Carotenoids were observed in **171 species among 65 genus** (we removed case where genus are mixed together, noted as "mixed"). We can see that *Diaptomus* and *Calanus* species are the most studied. 
</div>

<br/>
In some articles, carotenoid pigmentation was just discussed in the text of the article but not quantified, so we indicated "no_quantification" in the "study_type" column. **For following analyses, only studies that quantified carotenoids will be used:**

```{r results='hide'}
SI_obs_and_quanti <- SI
SI_obs_and_quanti %>% distinct(study, year, no, latitude, longitude) %>% nrow()
# -> pigmentation (red or blue) was observed at 154 locations globally
SI_obs_and_quanti %>% nrow()
# -> 316 observations

# Keep only quantitative measures
SI <- SI %>% filter(str_detect(study_type, "no_quantification", negate = TRUE))
SI %>% distinct(study, year, no, latitude, longitude) %>% nrow()
# -> 128 locations
SI %>% nrow()
# -> 286 observations
```



<div style="background-color: #bbe1e0">
Red or blue pigmentation of copepod was reported in **154 locations globally** for 316 observations. However, if we remove the 30 studies that reported or analysed pigmentation without a quantitative approach, we obtain **286 measurements of pigments**  (combinations study/latitude/longitude/altitude/species/stages) at 128 locations. <br/>

</div>

<br/>

### *Units*

```{r}
SI %>% group_by(unit, unit_conversion) %>% summarise(nb_of_studies_using_this_unit = n()) %>% arrange(-nb_of_studies_using_this_unit) %>% ungroup() 
#%>% write_xlsx("Summary_units.xlsx")

```

<div style="background-color: #bbe1e0">
Among the **286 quantifications the most common unit was µg/mgDW (202 measurements)**. 11 NA are present because for 5 studies, carotenoids were quantified to distinguish proportions of various pigments forms but no value of total carotenoid content was provided. 9 measurements have been successfully converted into µg/mgDW. We were not able to convert the other 64 values because of the method used (ex: mixed species or development stages, missing information, new methods not comparable). All details of conversions are given in Appendix 1: Table S1. Finally, **211 values in µg/mgDW** will be used for all following quantitative comparisons.
</div>
<br/>

### *Freshwater vs marine ecosystems*

```{r}
# Filter measurments in µg/mgDW
ug <- SI %>% filter (unit == "ug/mgDW")

# Transform to have numeric values 
ug$concentration <- as.numeric(ug$concentration)

# Total carotenoid average 
mean(ug$concentration)
# -> 1.22

# Count number of studies by ecosystems 
ug %>% group_by(ecosystem) %>% summarise(counts = n(), `median concentration` = median(concentration)) %>% ungroup()
```
 Global average carotenoid concentrationfrom 211 measurments is 1.22 µg.mgDW-1.

```{r fig.height=10, fig.width=16, fig.cap=c("Figure 1. Carotenoid concentrations in freshwater and marine copepods at the individual level across the world.")}

# Density distributions of pigment concentrations in freshwater and marine ecosystems: 
pigment_dd <- ug %>% 
  ggplot +
  #geom_histogram(aes(x = mean, fill = ecosystem, color = ecosystem), alpha =0.85)+
  geom_density(aes(x = concentration, fill = ecosystem, color = ecosystem), alpha =0.85)+
  #scale_x_continuous(trans='log10')+
  scale_x_log10(breaks = trans_breaks("log10",  function(x) 10^x),
             labels = trans_format("log10", math_format(10^.x))) +
  annotation_logticks(sides="b")+
  labs(x="Carotenoids concentration (μg/mgDW)",
       y="Density",
       fill= "Ecosystem")+
  theme_bw(base_size=22)+
  ylim(0,1.5)+
  theme(legend.position = "none", panel.grid.minor = element_blank())+
  scale_fill_manual(values=c("#7fc97f","#0868ac"))+
  scale_color_manual(values=c("#7fc97f","#0868ac"))+
  geom_vline(xintercept = median(filter(ug, ecosystem == "marine")$concentration), linetype="dotted", 
                color = "#0868ac", size=1)+
  geom_vline(xintercept = median(filter(ug, ecosystem == "freshwater")$concentration), linetype="dotted", 
                color = "#7fc97f", size=1)


# World map with locations of pigmented copepod observations: 

# Create a dataframe with locations of all studies (not only quantifications because observation is enough)
locations <- SI_obs_and_quanti %>% distinct(study, year, no, latitude, longitude, ecosystem, blue_color) %>% ungroup()


# Project the studies locations 
loc_proj <- st_as_sf(filter(locations, is.na(latitude) == FALSE), coords = c("longitude", "latitude"), 
    crs = 4032, agr = "constant")

# Find coastlines and borders
world <- ne_countries(scale = "medium", returnclass = "sf")

# Plot the map 
world_map <- ggplot(data = world) +
  geom_sf(size=0.05)+ #plot the world map 
  geom_sf(data = loc_proj, size = 2, alpha = 0.85, aes(color = ecosystem, shape= ecosystem))+ #project data points
  #geom_sf(data = filter(loc_proj, blue_color == "yes"), size = 0.8, alpha = 1, color= "#66FFFF")+ #project data points
  #geom_sf(data = filter(loc_proj, blue_color == "no"), size = 0.5, alpha = 1, color= "#CC0000")+ #project data points
  coord_sf(crs = "+proj=laea +lat_0=42 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs ")+ #spherical projection
  labs(color= "Ecosystem", shape = "Ecosystem")+
  scale_color_manual(values=c( "#7fc97f", "#0868ac"))+
  theme_minimal(base_size=20)+
  theme(legend.position = c(1.72, 0.8))

ggdraw(pigment_dd) +  draw_plot(world_map, width =0.75, height =0.75, 
        x = -0.04, y = 0.24)

#ggsave("map_and_density.pdf", width=18, height =11)

```


<div style="background-color: #bbe1e0">

The inset map shows the 154 locations where carotenoid-based pigmentation have been observerd (blue or red). Disctributions of carotenoids contents were done with the 214 values expressed in µg.mgDW-1. Fweeshwater copepods have an median carotenoid concentration of **2.243 µg.mgDW-1 (n=66)** and marine species a median carotenoid concentration of **0.137µg.mgDW-1 (n=145)**. 
</div>

<br/>
<br/>

We performed statisical tests to compare carotenoid mean concentrations between freshwater and marine ecosystems

```{r results='hide'}
ug$ecosystem = factor(ug$ecosystem, levels = c("freshwater", "marine"))

# First, test of normal distribution of data 
shapiro.test(filter(ug, ecosystem =="freshwater")$concentration)
# -> W = 0.76516, p-value = 6.513e-09
shapiro.test(filter(ug, ecosystem =="marine")$concentration)
# -> W = 0.39191, p-value < 2.2e-16

```
-> concentrations are not normally distributed

```{r results='hide'}
ug$concentration_log10 <- log10(ug$concentration)
# -> log transformation

shapiro.test(filter(ug, ecosystem =="freshwater")$concentration_log10)
# -> W = 0.98421, p-value = 0.5655 : normal distribution
shapiro.test(filter(ug, ecosystem =="marine")$concentration_log10)
# -> W = 0.9587, p-value = 0.0003215
# -> log-transformed marine concentrations are still not normal after transformation

leveneTest( log10(concentration) ~ ecosystem, data = ug, alternative = "two.sided" )
# Df F value   Pr(>F)   
#1  16.172 8.079e-05 ***

```
-> log-transformed marine concentrations were **still not normal, and variances are not equal**, so we will use a non-parametric test


```{r}
# So we can use a non-parametric test: 
wilcox.test(filter(ug, ecosystem =="marine")$concentration_log10, filter(ug, ecosystem =="freshwater")$concentration_log10, alternative = "two.sided")
# -> significantly different 

```

<div style="background-color: #bbe1e0">
Freshwater and marine carotenoid concentrations are **significantly different** using a Wilcoxon rank sum test. We can then observe mean concentration differences between copepod genus:  
</div>

```{r fig.align="center", fig.cap=c("Figure S1. Carotenoid concentration (µg/mgDW) in different freshwater and marine copepod genus. Color represent the number of studies that were used to compute the mean carotenoid content."), fig.height=11, fig.width=14}
ug %>% filter(genus != "mixed") %>% 
  group_by(genus, ecosystem) %>% 
  summarise(nb_studies = n(), pigment_content = mean(concentration), sd = sd(concentration)) %>% 
  ungroup() %>% 
  ggplot() + 
  geom_bar(aes(pigment_content, genus, fill = nb_studies), stat ="identity")+
  facet_wrap(~ecosystem)+
  labs(x="Carotenoid content (µg/mgDW)", fill = "Number 
of studies")+
  theme_bw(base_size = 16)

ggsave("fig_sup/Fig S1.png", width=10, height =10)
```


<br/>

<div style="background-color: #bbe1e0">
Plenty of copepod genus have the availability to synthesize astaxanthin but their propensity to do so seems mediated by the environment where they live. Note that then marine species accumulating the more pigments are *Tigriopus sp.* living in shallow temperate waters (ex: tidal pools of rocky shores) and *Calanus spp.* living in arctic and subarctic waters, submitted to strong seasonal climate variations.  
</div>

<br/>

## B. Localisation, molecular forms and metabolism of carotenoids pigments

### *Copepods' carotenoids types (astaxanthin, canthaxanthin, etc)*

```{r results='hide'}
# Number of studies that quantified the different types of carotenoids in copepods
detailed_carotenoids <- SI %>% filter(!is.na(astaxanthin_on_total_car)) 
detailed_carotenoids %>% group_by(study, year, no) %>% summarise(obs = n()) %>% nrow()
# -> 32 studies 

asta_car <- SI %>% group_by(astaxanthin_on_total_car) %>% summarise(counts = n())
sum(asta_car$counts) # on a total of 286 observations
asta_car <- asta_car %>% filter(!is.na(astaxanthin_on_total_car)) # remove NA
sum(asta_car$counts) 
# -> 170 observations with detailed information about carotenoids and 116 observations without detailed information about carotenoids

# For these 32 studies and over 170 quantification, distinguishing  observations of astaxanthin in minority (<50%) or majority(>50%)
  # Characters strings : 
min1 <- filter(asta_car, astaxanthin_on_total_car == "minority")
dom1 <- filter(asta_car, astaxanthin_on_total_car == "dominance")

  # Numerical values (remove characters)
asta_car <- asta_car %>% filter(astaxanthin_on_total_car!= "minority" & astaxanthin_on_total_car != "dominance")
asta_car$astaxanthin_on_total_car <- as.numeric(asta_car$astaxanthin_on_total_car)
dom2<- asta_car %>% filter(astaxanthin_on_total_car >= 50)
min2 <- asta_car %>% filter(astaxanthin_on_total_car < 50)

  # Bind into dataframes in order to count
dom <- rbind(dom1, dom2)
min <- rbind(min1, min2)
sum(dom$counts) / 170 *100
sum(min$counts) / 170 *100
# -> in 96%, astaxanthin was dominant (and really often largly dominant : >90% of copepod carotenoids) - 164 quantification 
# -> in ~4%, astaxanthin was in a minority or <50% of carotenoids - 6 quantifications 

# Canthaxanthin : 
asta_cantha <- SI %>% filter(str_detect(comments, "canthaxanthin")) %>% select(study, year, no, comments) # if authors mentioned canthaxanthin in one observation, we always reported it in the column "comments" of the dataset
setdiff(asta_cantha$study, detailed_carotenoids$study) # coherently, all studies that mentioned canthaxantin (9) are studies that have detailed all the carotenoids of copepods (30)
nrow(asta_cantha %>% group_by(study, year, no) %>% summarise(n()))
nrow(asta_cantha)
nrow(asta_cantha) / 170 *100
# -> canthaxanthin was found in only 9 studies (20 observations, 12% of cases), and almost everytime in "negligible amounts, traces or <5%"
```


<div style="background-color: #bbe1e0">
**Thirty-two studies** provide a detailed description of carotenoid pigments molecular forms for various species and developmental stages (170 measurements). In **96% (164 quantifications), astaxanthin was dominant** (or >50%), in ~4% (6 quantifications) of the cases, astaxanthin was in a minority or <50% of carotenoids. **Canthaxanthin was found in only 9 studies** (20 observations, 12% of cases), and almost every time in "negligible amounts, traces or <5%". 
</div>

<br/>

### *Astaxanthin forms*

```{r results='hide'}
# Number of studies that detailed astaxanthin forms (free, mono and di-esters)
nrow(SII) # 38 detailed observations of astaxanthin forms
distinct(SII, study, year, order) # 18 studies

# Character strings : 
free_dom_est_min1 <- SII %>% filter(free == "dominant" & all_esters == "minority")
nrow(free_dom_est_min1) # Free form dominant and esters in minority: 7 obs
free_min_est_dom1 <- SII %>% filter(free == "minority" & all_esters == "dominant")
nrow(free_min_est_dom1) # Ester form dominant and free form in minority: 3 obs

# Remove charcter strings to have only numerical values
SII <- SII %>% filter(free != "dominant" & free != "minority")
SII$free <- as.numeric(SII$free)
SII$all_esters <- as.numeric(SII$all_esters)
free_dom_est_min2 <- SII %>% filter(free >= 50)
nrow(free_dom_est_min2) # Free form >= 50%: 16obs
free_min_est_dom2 <- SII %>% filter(free < 50)
nrow(free_min_est_dom2) # Free form < 50%: 12obs

free_dom_est_min <- rbind(free_dom_est_min1, free_dom_est_min2)
free_min_est_dom <- rbind(free_min_est_dom1, free_min_est_dom2 )

nrow(free_dom_est_min) / 38 * 100  
# -> in 61% of cases, free astaxanthin is dominant
nrow(free_min_est_dom)  / 38 * 100
# -> in 39% of cases astaxanthin esters are dominant

# Look at the proportions of mono- and di-esters
rbind(free_dom_est_min2, free_min_est_dom2) %>% 
  filter(!is.na(`mono-esters`)) %>% 
  select(- all_esters) %>% 
  gather(key = "asta_form", value = "percent" , "free", `mono-esters`,`di-esters`) %>% 
  group_by(asta_form) %>% 
  summarise(mean_percent = mean(percent))
# -> very detailed analyses (8 quantification from 7 studies) show that on average, 53% of astaxanthin is in the free form, 27% of astaxanthin is in mono-esters and 20% is di-ester. 

```


<div style="background-color: #bbe1e0">
**18 studies** provided detailed forms of carotenoids (38 observations). In **61%** of cases, **free astaxanthin is dominant**, and in **39%** of cases **esters are dominant**. Very detailed analyses (8 quantifications from 7 studies) show that on average, 53% of astaxanthin is in the free form, 27% of astaxanthin is in mono-esters and 20% is di-ester.
</div>
<br/>

### *Ontogenic development*

This graphic illustrates that we don't have enough study that quantified carotenoid concentration along ontogenic development. We can not conclude about carotenoid quantities and forms according to copepod stages. <br/>

```{r fig.height=3, fig.width= 5, fig.align="center", fig.cap=c("Carotenoid concentrations in copepod stages from three different studies. Points are colored according to the unit used.")}
SI %>% filter((study == "Hairston" & no == "b")| study == "Holeton et al." | study == "Lotocka et al.") %>% mutate(Stage = factor(stage,levels = c("eggs", "nauplii", "copepodits", "adults", "adults_females"))) %>% ggplot() + geom_point(aes(Stage, concentration, color = unit, shape = study), size=3)+ labs(y="Carotenoid concentrations (units in color)", x="Development stage")+ theme_bw(base_size = 10)
```
<br/>
<br/>

### *Blue color*
```{r fig.height=3, fig.width= 5, fig.align="center", fig.cap=c("Latitudinal distribution of blue copepod observations")}
blue_cops <- SI_obs_and_quanti %>% filter(blue_color == "yes")
# ->  27 observations of blue copepods 

blue_cops %>% group_by(genus) %>% summarise(nb_obs = n()) %>%  arrange(-nb_obs)
# -> Labidocera and Pontella genus

# Latitudinal distribution
blue_cops %>% 
  filter(ecosystem=="marine") %>% 
  ggplot()+ 
  geom_density(aes(latitude))
```
<br/>
<br/>

<div style="background-color: #bbe1e0">
**27 observations of blue copepods**, all in marine ecosystems except for 1 freshwater case. Blue copepods mostly belong to Labidocera and Pontella genus. Most observations were made around 25°N, and none beyond 50°N. 

</div>
<br/>


# II. Effects of abiotic and biotic forcings on astaxanthin concentrations in copepods

## A. Environmental controls on pigmentation 


### *Altitude*

```{r}
test_marine <- ug %>% filter(ecosystem == "marine") %>% mutate(lat_abs= abs(latitude))
test_freshwater <- ug %>% filter(ecosystem == "freshwater") %>% mutate(lat_abs= abs(latitude))
```

```{r fig.align="center", fig.cap=c("Figure S2. Carotenoid concentration (µg/mgDW, log-transformed) according to lake altitude (m) with a linear model fitted (significant, p-value < 0.001)."), fig.height=3, fig.width=5}

ug  %>% filter(ecosystem=="freshwater") %>% ggplot() + 
  geom_point(aes(altitude, log10(concentration)), color="grey40") + 
  geom_smooth(aes(altitude, log10(concentration)), method = "lm", color="black")+
  labs(x="Lake altitude asl. (m)", y=" Carotenoid content (µg/mgDW) 
  Log-transformed values")+
  theme_bw(base_size = 10)

ggsave("fig_sup/Fig S2.png", width=6, height =4)
```


```{r, fig.align="center", fig.height=5, fig.width=6}
model <- lm(log10(test_freshwater$concentration)~test_freshwater$altitude)
summary(model)
par(mfrow=c(2,2)) 
plot(model)

```


<div style="background-color: #bbe1e0">
There is a **significant relationship between carotenoids content and altitude of lakes, p-value < 0.001**. However, we see a lot of noise (R square = 0.10) that could be explain by variations of other abiotic variables (temperature, lake depth, water turbidity) and biotic variables (phytoplanktonic community, predators, disponibility of other UV protectants). Extreme high value, but no reason to remove it, as it comes from a very shallow pond in high altitude.
</div>
<br/>

### *Latitude*

```{r}
summary(lm(log10(test_marine$concentration)~test_marine$lat_abs))
summary(lm(log10(test_freshwater$concentration)~test_freshwater$lat_abs))

```

```{r fig.align="center", fig.cap=c("Figure S3. Carotenoid concentration (µg/mgDW, log-transformed) according to latitude (° in absolute value) in freshwater and marine ecosystems. Linear models are fitted but are not significant."), fig.height=3, fig.width=5}
ug  %>% 
  ggplot() + 
  geom_point(aes(abs(latitude), log10(concentration), color = ecosystem))+ 
  geom_smooth(aes(abs(latitude), log10(concentration), group = ecosystem, color = ecosystem), method = lm)+
  scale_color_manual(values=c("#7fc97f","#0868ac"))+
  theme_bw(base_size = 10)+
  labs(y= "Carotenoid content (µg/mgDW)
  Log-transformed values", x="Latitude (°) in absolute value", color = "Ecosystem")

ggsave("fig_sup/Fig S3.png", width=7, height =4)

```


<div style="background-color: #bbe1e0">
**Carotenoids content shows no relationship with latitude, neither in freshwater nor in marine environments.** In fact, concentration variations within a study (or a species) are often more important than variations according to latitude. </br>
</div>
<br/>


## B. Biological and metabolic drivers of pigmentation

### *Seasons*

```{r}
SI %>% filter(is.na(seasonal_pic_season) == FALSE) %>% select(study, year, no, seasonal_pic_season) %>% group_by(seasonal_pic_season) %>% summarise(count = n()) %>% mutate(tot = sum(count), percent = count/tot*100) %>% select(-tot)
```

<div style="background-color: #bbe1e0">
On **37 time series**, astaxanthin content showed seasonal patterns with **maxima in winter (45%) or in spring (40%)** or in automn (3%). In 10% of the cases, no seasonal maxima were observed. 
</div>

<br/>


## C. Impacts of intra- and inter-species interactions on pigmentation

### *Predation in lakes*
```{r fig.align="center", fig.cap=c("Figure S4. Distribution of carotenoid concentrations (µg/mgDW, log-transformed) from lakes with and without predator presence. Carotenoid concentration difference according to predator presence is significant (Wilcoxon test, p-value = 0.03597)"), fig.height=3.5, fig.width=5.5}
ug$predation = ifelse(ug$predator_abundance=="null", "no", "yes")

ug  %>% filter(ecosystem=="freshwater") %>%  group_by(predation) %>% summarise(n())
# -> predator presence: 33 yes, 18 no

ug  %>% filter(ecosystem=="freshwater" & is.na(predation)== FALSE) %>% 
  ggplot() + 
  geom_boxplot(aes(predation, concentration)) + 
  geom_point(aes(predation, concentration, color=altitude)) +
  scale_color_viridis_c()+
  scale_y_continuous(trans='log10')+
  labs(y="Carotenoid contents (μg/mgDW) 
  Log10 axis transformation", color = "Altitude (m. asl)", x= "Fish presence")+
  theme_bw(base_size = 10)

ggsave("fig_sup/Fig S4.png", width=6, height =4)
```


```{r results='hide'}
#Test comparison predation/no predation (<30)
df_fish <- ug %>% filter(predation == "yes" | predation == "no") %>% mutate(predation = as.factor(predation))

#Normality ? 
shapiro.test(log10(filter(df_fish, predation =="yes")$concentration))
# W = 0.62321, p-value = 3.945e-08 -> not normal distribution
shapiro.test(log10(filter(df_fish, predation =="no")$concentration))
# W = 0.92556, p-value = 0.09914 -> normal distribution

```

Because n<30 and data not normaly distributed, we use a **non-parametric test** to compare carotenoid concentrations between lakes **with or without predators**:

```{r}
wilcox.test(concentration ~ predation, data = df_fish)
df_fish %>% group_by(predation) %>% summarise(n(), mean(concentration))
```


<div style="background-color: #bbe1e0">
There is a significant difference in carotenoid content **(p-value = 0.03597)** between lakes **with** (mean carotenoid concentration = **2.06 µg/mgDW, n=34**) and **without fish** (mean carotenoid concentration = **3.35 µg/mgDW, n = 22**). Altitude could play a role inside each category : high altitude lakes or ponds without fish can also be shallow and exposed to strong light radiations.
</div>

<br/>


### *Summary of all abiotic and biotic drivers*

```{r, fig.width=6, fig.height=8, fig.align='center', fig.cap=c("Figure 4. Counts of correlations between copepod carotenoid pigmentation and main environmental, biological or ecological forcings tested in the litterature.")}
# Change astaxanthin evolution column into factor
SIII$asta_correlation <- factor(SIII$asta_correlation, levels =c("positive","none", "negative"), ordered = TRUE)

# Replace names to have a more lisible figure 
SIII$forcing_variable <- SIII$forcing_variable %>% 
  str_replace_all(c("diet_composition" = "A. Diet quality", "feeding_activity" = "B. Feeding", "diet_quantity" = "C. Diet Quality", "fish_presence" = "D. Predator", "diurnal_variations" = "E. Daily cycle", "UV_radiations" = "F. UV", "natural_light" = "G. Natural light", "UV_fish" = "H. UV + Predator", "latitude_discrete" = "I. Latitude", "individual_depth" = "J. In situ depth", "ice_cover" = "K. Ice cover", "temperature" = "L. Temperature", "lake_elevation" = "M. Lake altitude", "dissolved_organic_carbon" = "N. DOC", "copper"= "O. Copper", "lake_depth" = "P. Lake depth", "egg_ratio"= "Q. Egg ratio", "body_length" = "R. Body length", "lipid_content" = "S. Lipid content", "MAAs" = "T. MAAs"))
  
  
# Count the number of times a forcing variables was statistically tested according to carotenoid concentrations through a positive, negative or absent correlation:
SIII_grouped <- SIII %>%
  group_by(forcing_variable) %>% 
  mutate(n=n()) %>% 
  filter(n>=2) %>% 
  ungroup() %>% 
  group_by(forcing_variable, asta_correlation) %>% 
  summarise(counts=n()) %>% 
  ungroup()
  
SIII_grouped %>% 
  filter(forcing_variable != "removal of UV radiations" & forcing_variable != "UV_temp" & forcing_variable != "dvm") %>% #because these 3 variables are rare and too difficult to interpret without context 
  ggplot() + 
  facet_wrap(~forcing_variable, ncol = 4)+
  geom_bar(aes(x=asta_correlation, y=counts, fill=asta_correlation), stat="identity")+
  labs(x = "Types of correlation between various forcing 
    variables and individual carotenoid content:", y= "Counts of tested correlations in the litterature", fill = "Correlation")+
  scale_fill_manual(values=c("#238b45", "#91bfdb", "#990000"))+
  theme_bw(base_size = 12)+
  theme(strip.background = element_rect(colour="black", fill="white", linetype="solid"),
        axis.text.x = element_blank(), legend.position="bottom", legend.box = "horizontal",
         legend.title = element_blank(), legend.text = element_text(size = 12))

```
<br/><br/>

# III. How does redness impact copepod fitness (reproduction, growth, survival) ?

```{r}
# If we keep the whole dataset : 
SIV %>% 
  group_by(fitness_effect) %>%
  summarise(n=n(), percent = n()/(33+13+9))
# 9 negative (16%), 13 neutral (24%) and 33 positive effects (60%). 

```


If we keep the whole dataset (SIV), we can find 9 negative (16%), 13 neutral (24%) and 33 positive effects (60%) of carotenoid pigmentation on copepod fitness. However, we decided to **remove data tagged with "low confidence"** (weak statistical tests or poor data quality discussed in the text of the article, see column "comments" of SIV for more details). 


<br/>



```{r}
# However, we can remove points with "low" confidence, (bad statistics or data quality discussed in the text)
SIV %>% filter(confidence != "low") %>% 
  group_by(fitness_effect) %>%
  summarise(n=n(), percent = n()/(31+11+8))
# 8 negative effects (16%), 11 neutral (22%), 31 positive effects (62%)

# If we look in particular to survival and mortality curves : 
SIV %>% filter(str_detect(response_variable, "Survival") | str_detect(response_variable, "Mortality")) %>% 
  group_by(fitness_effect) %>%
  summarise(n=n(), percent = n()/21)
```

<div style="background-color: #bbe1e0">
Out of **50 statistically tested** effects on one fitness variable, red pigmentation was **unfavorable for 8 (16%), neutral for 11 (22%), and favorable in 31 cases (62%)**. If we look in particular to survival and mortality curves, when copepods are in general submitted to an external stressor, we can see that carotenoid pigmentation is favorable in **71% of the cases**.
</div>

<br/>

### *Counts of redness effects on various fitness variables*

```{r, fig.width=18, fig.height=10, fig.align='center', fig.cap=c("Figure 5. Carotenoid pigmentation effects on reproduction, growth and survival of copepods.")}

# Counts by fitness effect, arrange by fitness component (growth, survival or reproduction)
SIV_grouped <- SIV %>%
  filter(response_variable != "NA")%>%
  filter(confidence!= "low") %>% 
  group_by(fitness_effect, response_variable, graph_order,fitness_component) %>%
  summarise(n=n()) %>% 
  ungroup()%>%  
  spread(key=fitness_effect, value= n) %>% 
  arrange(-graph_order)

# Create a factor to have variables in a good order on the graphic (according to fitness component and discussion order in the body of the article)
SIV_grouped <- SIV_grouped %>%  mutate(index= as.factor(seq(1:nrow(SIV_grouped))))

# Favorable ---
fav <- SIV_grouped %>% select(response_variable, favorable, index, fitness_component) #keep only postive to have a color scale by effect type
fav$favorable <- as.factor(fav$favorable) # transform as factor to have the best color scale
green <- c("#41ae76", "#238b45", "#006d2c","#00441b") #because 4 levels

pfav <- ggplot(fav, aes(y = index, x= "", fill = favorable))+
  geom_tile()+
  scale_fill_manual(values=green, na.value = "white", na.translate = F)+
  scale_y_discrete(labels=fav$response_variable)+
  labs(y="Fitness response variables", title = "Favorable - 31 cases (62%)")+
  theme_minimal(base_size = 15)+
  theme(legend.position="bottom",   axis.title.x = element_blank(), legend.title = element_blank(), legend.text = element_text(size = 15))

# Neutral ---
neu<- SIV_grouped %>% select(response_variable, neutral, index) #keep only postive to have a color scale by effect type
neu$neutral<- as.factor(neu$neutral) # transform as factor to have the best color scale
blue <- c("#a6bddb", "#74a9cf", "#3690c0") #because 3 levels

pneu <- ggplot(neu, aes(y = index, x = "", fill = neutral))+
  geom_tile()+
  scale_fill_manual(values=blue, na.value = "white", na.translate = F)+
  theme_minimal(base_size = 15)+
  labs(x= "Redness effects on copepod fitness - counts of tested observations:", title = "Neutral - 11 cases (22%)")+
  scale_y_discrete(labels=element_blank())+
  theme(legend.position="bottom", legend.title = element_blank(),  axis.title.y = element_blank(), legend.text = element_text(size = 15))

# Unfavorable ---
unfav<- SIV_grouped %>% select(response_variable, unfavorable, index) #keep only positive to have a color scale by effect type
unfav$unfavorable <- as.factor(unfav$unfavorable) # transform as factor to have the best color scale
red <- c("#ef3b2c", "#cb181d","#a50f15") #because 3 levels

punfav <- ggplot(unfav, aes(y = index, x = "", fill = unfavorable))+
  geom_tile()+
  scale_fill_manual(values=red, na.value = "white", na.translate = F)+
  theme_minimal(base_size = 16)+
  scale_y_discrete(labels=element_blank())+
  labs(fill = "", title = "Unfavorable - 8 negative cases (16%)") +
  theme(legend.position="bottom", axis.title.y = element_blank(), axis.title.x = element_blank(), legend.title = element_blank(), legend.text = element_text(size = 18))


pfav | pneu | punfav
```







